{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-09 Paper/main.tex
part two proofreading before switching to pdf

	\subsection{Motivation}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC could be uncovered. An hypothesis for this is that the dependency between the truncation parameter $L$ and the FPC's, which are calculated from $L$-truncated smoothed curves, is too complex to capture in our setting. To examine this further, additional simulations ( section \ref{Tables_sim_additional}) were conducted for B-spline basis functions with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$. As in the simulation study we can find potential signs of overfitting for all setups and both choices of $L$. The additional simulations revealed evidence that there might exist a relationship from $n_{FPC} = 3$ onwards: For $n_{FPC} \in \{3, \dots, 7 \}$, the 10-fold cross-validated MSPE is always lower for 50 than for 70 basis functions which might be a first sign of a too high $L$ and subsequent undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, the FPCR performs better using a more conservative number of basis functions (here $L = 50$). The settings $f_1,Y_2$ and $f_2,Y_2$ perform worse for these high number of basis functions compared to the main simulation, while $f_1,Y_1$ and $f_2,Y_1$ show a stronger performance than in the main simulation study, although the signal-to-noise ratios of the responses are not affecting the construction of the FPC's. This might indicate beneficial attributes of the lower truncation parameters in settings with low signal-to-noise responses.

@@

	\subsection{Interpretation of Results}
		\subsubsection{Basis Expansion Regression}
		The cross-validated MSPE selects 5 basis functions for the Monomial basis. For B-Splines, the cross-validation selects 10 basis functions and 9 basis functions for the Fourier basis. Comparing the estimates in Appendix \ref{Estimates_Appl}, away from the boundaries, the B-spline and Fourier bases show similar behavior, opposing to the Monomial estimate which might be attributable to the lower number of basis functions. Especially at the lower boundary, the Monomial basis shows an exaggerated behavior. However, we must exercise caution since from $L=6$ onwards, we were not able to calculate stable results for the Monomial basis. The MSPE for B-splines (0.04574) and Fourier (0.04808) are similar while the Monomial basis (0.24181) shows distinctly worse properties. In contrast to the simulation study, the B-splines basis manages to outperform the Fourier basis. Driving factors for this better performance could be that, first, we do not assume a periodic true coefficient function with the same start- and end value and second, that the Fourier basis enforces identical start and end value on $NIR$. It is worthy to note that the reported MSPE for the B-spline and Fourier basis are close to the errors reported in the simulation study for the setup $f_2,Y_1$, which might be caused by the hypothesized similarities between the true coefficient function and $f_2$, but also by a similar high signal-to-noise ratio of the octane numbers.

@@

	\paragraph{Low Number of Basis Functions in FPCR Regression}
	Another limitation of the exploration into functional principal component regression is the low number of principal component specifications used in this paper. As the simulation and application showed, the number of basis functions does not seem to be the deciding factor for the performance of this method as long as an appropriately high number of basis functions is provided. 
	A more significant factor seems to be the number of principal components available for the linear regression which in itself could be subject to choice according to cross validation.
	\vspace{-0.2cm}

@@

	\paragraph{Orthogonal Polynomials to Solve Collinearity Problems of Monomial Basis}
	To address the collinearity problems described earlier, one possible idea would be to use a system of pair-wise orthogonal polynomials as a basis instead of the Monomial basis. One example is the system of Legendre polynomials which are orthogonal by construction and have the same closed span as the monomials. Due to their orthogonality, the problem of collinearity in the regressor matrix are greatly reduced which could allow for larger numbers of basis functions in the basis expansion regression approach. The first eight Legendre polynomials are shown in Appendix section \ref{Basis_Plots}.
	\vspace{-0.2cm}

@@

	\paragraph{Larger Range for the Number of Specifications for FPCR}
	As alluded to before, the chosen specifications for the number of principal components and the number of basis functions that are used for FPCR might be insufficient to obtain a full picture of the performance of this procedure when combined with smoothing using basis truncation. It would therefore be interesting to focus more attention on this aspect of the paper in a future extension to explore a broader array of possible specifications which might allow the derivation of more detailed insights.
	\vspace{-0.2cm}

	\paragraph{Input from Physics}
	It would be interesting to combine the purely mathematical findings of this paper with theoretical expertise from the field of physics as this could inform the choice of models and be conducive to a more meaningful interpretation of estimates. For example, specific parts of the NIR spectrum can possibly be liked to specific molecules associated with a high predictive power for the octane value. This could guide, for example, the construction of a B-spline basis that focuses on the relevant parts of the spectrum by using a vector of knots founded by field expertise. Alternatively, it could be used to link specific principal components to compounds which would be interesting for the interpretation of estimates.
	\vspace{-0.2cm}

>>>>

	\subsection{Motivation}
	\label{sim_motivation}

@@

	This convex behavior was not observed for the FPCR where no clear relationship between the basis functions and the number of FPC could be uncovered. An hypothesis for this is that the dependency between the truncation parameter $L$ and the FPC's, which are calculated from $L$-truncated smoothed curves, is too complex to capture in our setting. To examine this further, additional simulations ( Appendix \ref{Tables_sim_additional}) were conducted for B-spline basis functions with a large truncation parameter $L \in \{50, 70\}$ for $n_{FPC} \in \{2, \dots, 7 \}$. As in the simulation study we can find potential signs of overfitting for all setups and both choices of $L$. The additional simulations revealed evidence that there might exist a relationship from $n_{FPC} = 3$ onwards: For $n_{FPC} \in \{3, \dots, 7 \}$, the 10-fold cross-validated MSPE is always lower for 50 than for 70 basis functions which might be a first sign of a too high $L$ and subsequent undersmoothing. Following this, it seems that once a sufficient number of basis functions is used to expand the curves, the FPCR performs better using a more conservative number of basis functions (here $L = 50$). The settings $f_1,Y_2$ and $f_2,Y_2$ perform worse for these high number of basis functions compared to the main simulation, while $f_1,Y_1$ and $f_2,Y_1$ show a stronger performance than in the main simulation study, although the signal-to-noise ratios of the responses are not affecting the construction of the FPC's. This might indicate beneficial attributes of a lower truncation parameter in settings with low signal-to-noise responses.

@@

	\subsection{Interpretation of Results}
		\subsubsection{Basis Expansion Regression}
		The cross-validated MSPE selects 5 basis functions for the Monomial basis. For B-Splines, the cross-validation selects 10 basis functions and 9 basis functions for the Fourier basis. Comparing $\hat{\beta}(t)$ for the different bases in figure \ref{basis_expansion_estimate_appl}, away from the boundaries, the B-spline and Fourier bases show similar behavior, opposing to the Monomial estimate which might be attributable to the lower number of basis functions. Especially at the lower boundary, the Monomial basis shows an exaggerated behavior. However, we must exercise caution since from $L=6$ onwards, we were not able to calculate stable results for the Monomial basis. The MSPE for B-splines (0.04574) and Fourier (0.04808) are similar while the Monomial basis (0.24181) shows distinctly worse properties. In contrast to the simulation study, the B-splines basis manages to outperform the Fourier basis. Driving factors for this better performance could be that, first, we do not assume a periodic true coefficient function with the same start- and end value and second, that the Fourier basis enforces identical start and end value on $NIR$. It is worthy to note that the reported MSPE for the B-spline and Fourier basis are close to the errors reported in the simulation study for the setup $f_2,Y_1$, which might be caused by the hypothesized similarities between the true coefficient function and $f_2$, but also by a similar high signal-to-noise ratio of the octane numbers.

@@

	\paragraph{Low Number of Basis Functions in FPCR Regression}
	As the simulation and application showed, the number of basis functions does not seem to be the deciding factor for the performance of FPCR, apart from the additional simulations which reported signs of undersmoothing for larger values of $L$. A more significant factor seems to be the number of FPC's available for the linear regression. This paper selects these components in order of their explanatory power of the variability around the mean function. For reasons outlined in Section \ref{sim_motivation} it might be better to make the selection of the FPC´s subject to some criteria such as cross validation or base their choice on insights and expertise from the respective field.
	\vspace{-0.2cm}

@@

	\paragraph{Orthogonal Polynomials to Solve Collinearity Problems of Monomial Basis}
	To address the collinearity problems described earlier, one possible idea would be to use a system of pair-wise orthogonal polynomials as a basis instead of the Monomial basis. One example is the system of Legendre polynomials which are orthogonal by construction and have the same closed span as the monomials (cf. \citep{Dattoli_2001}). Due to their orthogonality, the problem of collinearity in the regressor matrix are greatly reduced which could allow for larger numbers of basis functions in the basis expansion regression approach. The first eight Legendre polynomials are shown in Appendix section \ref{Basis_Plots}.
	\vspace{-0.2cm}

@@

	\paragraph{Larger Range for the Number of Specifications for FPCR}
	As alluded to before, the chosen specifications for the number of principal components and the number of basis functions that are used for FPCR might be insufficient and limited to obtain a full picture of the performance of this procedure when combined with smoothing using basis truncation. It would therefore be interesting to focus more attention on this limitation of this paper in a future extension to explore a broader array of possible specifications which might allow the derivation of more detailed insights.
	\vspace{-0.2cm}

	\paragraph{Input from Physics}
	It would be interesting to combine the purely mathematical findings of this paper with theoretical expertise from the field of physics as this could inform the choice of models, and be conducive to a more meaningful interpretation of estimates. For example, specific parts of the NIR spectrum can possibly be liked to specific molecules associated with a high predictive power for the octane value. This could guide, for example, the construction of a B-spline basis that focuses on the relevant parts of the spectrum by using a vector of knots founded by field expertise. Alternatively, it could be used to link specific principal components to chemical compounds which would be useful for the model selection and interesting for the interpretation of estimates.
	\vspace{-0.2cm}
